---
title: "Example Data Analysis Using Gapminder Dataset"
format: html
bibliography: references.bib
---

# Example Data Analysis

For this analysis, I am using the **Gapminder** dataset, which contains data about life expectancy, GDP per capita, and population across different countries and continents over time.

## Research Question
How does **life expectancy** vary with **GDP per capita** across different continents?

## Intended Audience
This analysis is intended for **policy makers**, **public health researchers**, and **general public** who are interested in understanding the impact of economic factors on health outcomes.

## Data Source
The dataset comes from the **Gapminder Foundation** and is available through the **gapminder** R package.
- Original dataset: [Gapminder Dataset](https://www.gapminder.org/data/)
- **Data Dictionary**:
  - \`country\`: Name of the country
  - \`continent\`: Continent where the country is located
  - \`year\`: Year of observation
  - \`lifeExp\`: Life expectancy at birth (in years)
  - \`gdpPercap\`: GDP per capita (in USD)
  - \`pop\`: Population size
  
## Load and Wrangle Data

```{r echo=FALSE}
library(gapminder)
library(dplyr)

# Load data
data(gapminder)

# Wrangle data: Select relevant columns, filter, and create new variables
gapminder_clean <- gapminder %>%
  select(country, continent, year, lifeExp, gdpPercap, pop) %>%
  filter(year == 2007) %>%
  mutate(gdp_log = log(gdpPercap)) %>%
  arrange(desc(lifeExp))

gapminder_clean

```

## Data Visualization

```{r echo=FALSE}
library(ggplot2)

# Plot 1: Life Expectancy by Continent
ggplot(gapminder_clean, aes(x = continent, y = lifeExp, fill = continent)) +
  geom_boxplot() +
  labs(
    title = "Life Expectancy by Continent in 2007",
    x = "Continent",
    y = "Life Expectancy (Years)",
    caption = "Source: Gapminder Dataset"
  )

# Plot 2: GDP per Capita vs Life Expectancy
ggplot(gapminder_clean, aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "GDP per Capita vs Life Expectancy",
    subtitle = "Relationship between economic status and health outcomes",
    x = "GDP per Capita (USD)",
    y = "Life Expectancy (Years)"
  )

# Plot 3: Life Expectancy by Population Size with Faceting
ggplot(gapminder_clean, aes(x = pop, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  facet_wrap(~ continent) +
  labs(
    title = "Life Expectancy by Population Size Across Continents",
    x = "Population Size",
    y = "Life Expectancy (Years)"
  )
```
## Callout Blocks
> **Note**: Higher GDP per capita is often associated with increased life expectancy, indicating the importance of economic stability for health.

> **Tip**: Policy interventions that focus on economic growth may also contribute to improved health outcomes.

## Summary
The analysis of the Gapminder data reveals that **life expectancy** tends to be higher in continents with higher average **GDP per capita**. Additionally, countries with larger populations tend to exhibit a diverse range of life expectancies. This analysis highlights the importance of both **economic** and **demographic** factors in determining health outcomes.

## Margin Content
::: {.callout-tip}
### Health and Wealth
Countries with higher GDP per capita generally enjoy better health outcomes, emphasizing the importance of economic development.
:::

## References
- Gapminder Foundation. "Gapminder Data." [Link](https://www.gapminder.org/data/)
- Rosling, Hans. "Factfulness: Ten Reasons We're Wrong About the World." 2018.
- OECD. "Health at a Glance 2019: OECD Indicators." [Link](https://www.oecd.org/health/health-at-a-glance/)

## Functions Used
### dplyr
- `select()`, `filter()`, `mutate()`, `arrange()`
### ggplot2
- `geom_boxplot()`, `geom_point()`, `geom_smooth()`, `facet_wrap()`
---
